### 2015-04-16 implementation-specs.tex
update specs.

\noindent
If the user cares to specify the desired verson of {\it mpileaks}, she can do so
by supplying {\tt @<version>} (line 2). Versions can be precise (e.g. {\tt @4.5.1}) or 
they can be ambiguous (e.g., @

To specify the compiler (line 3), the user
simply adds {\tt \%} followed by its name, and if a version {\it of the compiler} 
is desired, the same {\tt @} syntax can be used.  Here, {\tt gcc} refers to the full
gcc compiler suite: C, C++, Fortran 77, and Fortran 90 compilers.

To handle build-time options like debug flags and optional components, spack
allows named flags, or {\it variants} to be associated with packages.  
In the example (line 4), {\tt mpileaks} can be built with a {\tt debug} option by
adding {\tt +debug}.  The option can be similarly disabled with {\tt -debug}.  
This allows package authors to provide package-specific build options without 
having to extend Spack.  We use names for these to simplify the spec
syntax, and to prevent Spack's configuration space from becoming too fine-grained.
It would be violate our goal of conciseness, for example, to include detailed 
compiler flags in the spec syntax, but known sets of flags can be exported using
named variants.

To support cross-compilation, Spack includes the platform in the package spec (line 5).
Platforms begin with {\tt =} and take names like {\tt linux-ppc64} or {\tt bbq}.  They are
specified per-package; this allows front-end tools to depend on their back-end measurement
libraries with a {\it different} architecture on cross-compiled machines.

Lines 6 and 7 in the table show the key feature that enables Spack's flexibility.
The user can specify all of the above information not only for the package being
installed, but also for its {\it dependencies}.  To do this, the user needs only supply 
\verb|^| and the dependency's name.  If we need to build a new version with a specific
version of {\tt mvapich2}, we can simply add, e.g., \verb|^mvapich2@1.9|
to the spec, and it will build with that MPI version instead of the default.
This can also be done for multiple libraries in the same spec (line 7).  
As mentioned, Spack guarantees that there is only one version of any package in 
the spec DAG.  So, within the same DAG, each dependency can be uniquely identified by 
only its package name.  This frees the user from having to worry about DAG connectivity;
the user only needs to know that a package depends, somehow, indirectly on {\tt mvapich}
to add this constrain
uniquely identifies a dependency, and dependency constraints can be specified in any
 order.

@@

In our experience, however, users
frequently want bleeding-edge versions, and package managers frequently lag 
behind the latest releases.  Spack therefore has a capability to extrapolate
the latest URL, using the {\tt url} attribute as a model\footnote{This works
for consistently named URLs, and it can be specified per-version for packages
with inconsistent naming.}.
The user can ask for an unknown version on the command line and still
successfully download and install it.  Spack can also uses the same model to
scrape webpages and find new versions.

>>>>

If the user wants more specificity, {\tt mpileaks} can be augmented with additional
qualifiers.  Spack install will handle selecting a build configuration that satisfies
the user's request.  We describe additional qualifiers below.

{\bf Versions.}
Versions are specified by supplying {\tt @<version>} after the package name (line 2). 
Versions can be precise (e.g. {\tt @4.5.1}) or 
they can be ambiguous (e.g., @4) or a colon-separated, inclusive range (e.g., @4.4:4.5.1),
and Spack will pick an available compiler from the specified range.
The package in Figure~\ref{fig:mpileaks} lists two ``safe'' versions with checksums, but
in our experience users frequently want bleeding-edge versions.  Package managers
frequently lag behind the latest releases. 
Spack has a capability to extrapolate URLs from versions,
using the package's {\tt url} attribute as a model\footnote{This works
for packages with consistently named URLs}.  The user can request a specific
version on the command line, even if it is unknown to Spack,
and Spack will attempt to install it.  Spack also uses the same
model to scrape webpages and find new versions as they become available.

{\bf Compilers.}
To specify the compiler (line 3), the user
simply adds {\tt \%} followed by its name, and if a version {\it of the compiler} 
is desired, the same {\tt @} syntax can be used.  Here, {\tt gcc} refers to the full
gcc compiler suite: C, C++, Fortran 77, and Fortran 90 compilers.  Spack can auto-detect
compiler toolchains if they are in the user's PATH or they can be registered manually
through a configuration file.

{\bf Variants.}
To handle build-time options like debug flags and optional components, spack
allows named flags, or {\it variants} to be associated with packages.  
In the example (line 4), {\tt mpileaks} can be built with a {\tt debug} option by
adding {\tt +debug}.  The option can be similarly disabled with {\tt -debug}.  
This allows package authors to provide package-specific build options without 
having to extend Spack.  We use names for these to simplify the spec
syntax, and to prevent Spack's configuration space from becoming too fine-grained.
It would be violate our goal of conciseness, for example, to include detailed 
compiler flags in the spec syntax, but known sets of flags can be exported using
named variants.

{\bf Cross-compileation.}
To support cross-compilation, Spack includes the platform in the package spec (line 5).
Platforms begin with {\tt =} and take names like {\tt linux-ppc64} or {\tt bbq}.  They are
specified per-package; this allows front-end tools to depend on their back-end measurement
libraries with a {\it different} architecture on cross-compiled machines.

{\bf Dependencies.}
Lines 6 and 7 in the table show the key feature that enables Spack's flexibility.
The user can specify all of the above information not only for the package being
installed, but also for its {\it dependencies}.  To do this, the user needs only supply 
\verb|^| and the dependency's name.  If we need to build a new version with a specific
version of {\tt mvapich2}, we can simply add, e.g., \verb|^mvapich2@1.9|
to the spec, and it will build with that MPI version instead of the default.
This can also be done for multiple libraries in the same spec (line 7).  
As mentioned, Spack guarantees that there is only one version of any package in 
the spec DAG.  So, within the same DAG, each dependency can be uniquely identified by 
only its package name.  This frees the user from having to worry about DAG connectivity;
the user only needs to know that a package depends, somehow, indirectly on {\tt mvapich}
to add this constrain
uniquely identifies a dependency, and dependency constraints can be specified in any
order.

@@
